% Author: Mudit Gupta
% Email: mudit+csm16a@berkeley.edu
\learning{Prereqs: What are vector spaces and subspaces? \\
Description: Explains how to read set notation, tries to make students really realize that the notation means a set of vectors, and that a subspace is also a set of vectors. And what a subspace intuitively means.}

\begin{enumerate}

\item{
    Consider the set $W = \left\{\begin{bmatrix} a_1 \\ a_2 \\ a_3 \end{bmatrix}, a_1, a_2, a_3 \in \mathbb{R}: a_1 + 2a_2 - 3a_3 = 0\right\}$. Is $\begin{bmatrix} 1 \\ 2 \\ 3 \end{bmatrix}$ an element of the set $W$?
}

\answerbox{3cm}



\sol{
    The way set notation works is that for an element to be a part of the set, $a_1 + 2a_2 - 3a_3$ must be satisfied. We can plug in numbers for $a_1, a_2, a_3$ from the vector and see if the equation is satisfied. $1 + 4 - 9 \neq 0$, so this element is not a member of the set.
}

\item{Write any 3 elements from this set.}

\answerbox{3cm}

\note{The purpose here is really to make sure the students realize that $W$ is indeed a set. It has infinite number of elements, but it is still a set. Let the students come up with whatever they want. Ultimately though, it will be super helpful if the final 3 elements you get are the same as the ones in the answer below. We will be using these again! So verify a couple of the elements that the students put forth, but ultimately write these on the board.}

\sol{There are many possibilities. For instance, we could set the values of $a_2$ and $a_3$ to be $0$ and then see what value of $a_1$ works. $\vec{v}_1 = \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}$ is in the set. \\

Another element can be obtained by setting the values of $a_1$ and $a_2$ to be $1$, and then we get $1 + 2 - 3a_3 = 0$, or $a_3 = 1$. Therefore $\vec{v}_2 = \begin{bmatrix} 1 \\ 1 \\ 1 \end{bmatrix}$ is also in the set. \\

Another element can be obtained by setting the values of $a_1 = 3$ and $a_3 = 2$ and then we get $3 + 2a_2 - 6 = 0 \implies a_2 = \frac{3}{2}$. Therefore $\vec{v}_3 = \begin{bmatrix} 3 \\ \frac{3}{2} \\ 2 \end{bmatrix}$ is in the set too.}

\item{Is the set $W$ a subspace?}

\answerbox{6cm}

\note{
Meta:
If students attempt an exhaustive proof of all the axioms of a vector space, be sure to note that many are satisfied by nature of the superset R3. Explain what closure under (+, *) means and why it matters.
A fast way to disprove a subspace is to show the non-existence of the zero vector. 
Explain how set notation is defined and how to read it
Subset != subspace
}

\sol{
    \textbf{Step 1: Claim that $W$ is a sub$\textit{set}$ of, say, $X$}. \\
    $W$ is clearly a subset of $\mathbb{R}^3$. This can be seen because the elements of $W$ contain 3 elements, but $W$ is not equal to $\mathbb{R}^3$ since some elements from $\mathbb{R}^3$ are not in $W$. \\ 

    \textbf{Step 2: Claim that $X$ is a vector space} \\
    $\mathbb{R}^3$ is a vector space that we have seen in lecture. \\ 

    \textbf{Step 3: If $X$ is a known vector space, and $W$ is a subset of $X$, only 3 axioms must be proven.} \\ \\
    \textbf{a: Prove closure under addition} \\
    Consider two arbitrary elements from the set $\begin{bmatrix} b_1 \\ b_2 \\ b_3 \end{bmatrix}$ and $\begin{bmatrix} c_1 \\ c_2 \\ c_3 \end{bmatrix}$. Since these elements are a part of the set, it is true that $b_1 + 2b_2 - 3b_3 = 0$ and that $c_1 + 2c_2 - 3c_3 = 0$. \\
    Consider the sum of these elements. $\begin{bmatrix} b_1 + c_1 \\ b_2 + c_2 \\ b_3 + c_3 \end{bmatrix}$. Is this element a part of the set too? In other words, is $(b_1+c_1) + 2(b_2+c_2) - 3(b_3 + c_3) = 0$? \\
    $$(b_1+c_1) + 2(b_2+c_2) - 3(b_3 + c_3) \stackrel{?}{=} 0$$
    $$\implies (b_1 + 2b_2 - 3b_3) + (c_1 + 2c_2 - 3c_3) \stackrel{?}{=} 0$$
    Clearly, the left hand side equals the right hand side. Therefore, 
    $$0 + 0 = 0$$.
    Thus, the element $\begin{bmatrix} b_1 + c_1 \\ b_2 + c_2 \\ b_3 + c_3 \end{bmatrix}$ is a part of the set and we have proven closure under addition. \\ 

    \textbf{b: Prove closure under scalar multiplication} \\
    Consider an arbitrary element from the set $\begin{bmatrix} d_1 \\ d_2 \\ d_3 \end{bmatrix}$. This means that $d_1 + 2d_2 - 3d_3 = 0$ is true. \\
    Consider some scalar $s$. Is $s\begin{bmatrix} d_1 \\ d_2 \\ d_3 \end{bmatrix}$ in the set $W$? I.e., is $\begin{bmatrix} sd_1 \\ sd_2 \\ sd_3 \end{bmatrix}$ in the set $W$? I.e. is $sd_1 + 2sd_2 - 3sd_3 = 0$?
    $$sd_1 + 2sd_2 - 3sd_3 \stackrel{?}{=} 0$$
    $$\implies s \cdot (d_1 + 2d_2 - 3d_3) \stackrel{?}{=} 0$$
    Indeed the left hand side of the equation equals the right hand side, i.e., $$s\cdot 0 = 0$$

    Therefore, $\begin{bmatrix} sd_1 \\ sd_2 \\ sd_3 \end{bmatrix}$ is in the set $W$ and the set $W$ is closed under scalar multiplication. \\ 

    \textbf{c: Prove existence of 0 element} \\
    We need to check whether $\vec{0} = \begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}$ exists in the set. This is easy to check because we just need to check whether $0 + 2 \cdot 0 - 3 \cdot 0 = 0$, which it is. So the 0 element exists in the set. \\ 

    Therefore, the set $W$ is a subspace of $\mathbb{R}^3$. 
}

\item{How can we now quickly find more elements of this set?}

\answerbox{3.5cm}

\sol{
    Since we know the set is closed under scalar multiplication and under addition, we can easily find more elements. Previously, we found $\vec{v}_1 = \begin{bmatrix} 1 \\ 1 \\ 1 \end{bmatrix}$ to be an element. Now we know that any multiple of this is in the set too! \\
    We also found that $\vec{v}_2 = \begin{bmatrix} 3 \\ \frac{3}{2} \\ 2 \end{bmatrix}$ was in the set. Now we can add the 2 elements we found, and $\begin{bmatrix} 4 \\ \frac{5}{2} \\ 3 \end{bmatrix}$ is also in the set. In fact, any $s\vec{v}_1 + r\vec{v}_2$, where $s, r \in \mathbb{R}$ are in the set! 
}

\item{Consider the set $X = \left\{\begin{bmatrix} a_1 \\ a_2 \\ a_3 \end{bmatrix}, a_1, a_2, a_3 \in \mathbb{R}: a_1 * a_2 * a_3 = 0\right\}$. Is $X$? is subspace?}

\answerbox{4cm}


\sol{
    We can quickly see that this set would have a zero vector. \\

    This set also satisfies the scalar multiplication property; if $a_1 * a_2 * a_3 = 0$, one of the elements of the vector has to be zero and scaling the vector will not change that. \\

    However, this set is not closed under addition. We can verify this: \\

    $\begin{bmatrix} 1 \\ 1 \\ 0 \end{bmatrix}$ belongs to the set as $1 * 1 * 0 = 0$. \\

    $\begin{bmatrix} 0 \\ 0 \\ 1 \end{bmatrix}$ also belongs to the set as $0 * 0 * 1 = 0$. \\

    However, the sum of the two vectors, $\begin{bmatrix} 1 \\ 1 \\ 1 \end{bmatrix}$ doesn't belong to the set as $1 * 1 * 1 = 1$. \\

    Hence, set X is not a subspace.
}



\end{enumerate}